19Q1

### 2019-04-23 19Q1.Rmd

MergeOmicsG2P <- function(database, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  filt <- auc %>% filter(Database == database) %>% select(Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID, Drug, AUC_IC50)
  print(paste0("Filtered for ", database, "."))

@@

  # Merge mutation calls
  data <- merge(data, maf_df_filt, by = c("Gene", "Broad_ID"), all.x = TRUE)
  data$Mutation_Status <- ifelse(is.na(data$Mutation_Status), 0, data$Mutation_Status)
  ## Identify point mutations
  data$PointMutation <- ifelse(is.na(data$PointMutation), 0, data$PointMutation)
  ## Add color columns for plotting
  if(database == "CCLE") {
    data$Color <- ifelse(data$Mutation_Status == 0, "palegreen3", "springgreen4")
    data$Color <- factor(data$Color)
  }
  if(database == "CTRP") {
    data$Color <- ifelse(data$Mutation_Status == 0, "slategray3", "steelblue4")
    data$Color <- factor(data$Color)
  }
  if(database == "GDSC") {
    data$Color <- ifelse(data$Mutation_Status == 0, "paleturquoise3", "turquoise4")
    data$Color <- factor(data$Color)
  }
  print("Merged mutation calls.")

@@

@@

FilterForPointMuts <- function(dataframe, outfile) {
  # Filter for point mutations
  ptmuts <- filter(dataframe, PointMutation == 1)
  # Remove columns that lead to duplicate mutation entries
  ptmuts <- subset(ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, PointMutation))
  ptmuts$Gene <- ordered(ptmuts$Gene, levels = g2p_genes)
  ptmuts <- unique(ptmuts)
  saveRDS(ptmuts, outfile, compress = "xz")
  return(ptmuts)
}
```

@@

# ccle <- MergeOmicsG2P(database = "CCLE", outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
# ctrp <- MergeOmicsG2P(database = "CTRP", outfile = "./data_munging/rds/ctrp_data_19Q1_g2pgenes.rds")
# gdsc <- MergeOmicsG2P(database = "GDSC", outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")

# ccle_ptmuts <- FilterForPointMuts(dataframe = ccle, outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
# ctrp_ptmuts <- FilterForPointMuts(dataframe = ctrp, outfile = "./data_munging/rds/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- FilterForPointMuts(dataframe = gdsc, outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")
```

@@

```{r, eval=FALSE}
ccle <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

# Testing
test_ccle <- ccle_ptmuts %>% group_by(Gene, Mutation_Status_Nonsilent) %>% tally()
test_ctrp <- ctrp_ptmuts %>% group_by(Gene, Mutation_Status_Nonsilent) %>% tally()
test_gdsc <- gdsc_ptmuts %>% group_by(Gene, Mutation_Status_Nonsilent) %>% tally()
```

@@

```{r, eval=FALSE}
crispr <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
colnames(crispr)[1] <- "Broad_ID"
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Gene", value.name = "Score")
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_meta, by = "Broad_ID", all.x = TRUE)
# Merge mutation annotations
maf_filt_crispr <- filter(maf_df, Gene %in% as.character(unique(crispr_melt$Gene)))
maf_filt_crispr <- subset(maf_filt_crispr, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1))
maf_filt_crispr <- unique(maf_filt_crispr)
crispr_muts <- merge(crispr_melt, maf_filt_crispr, by = c("Gene","Broad_ID"), all.x = TRUE)
crispr_muts$Gene <- factor(crispr_muts$Gene)
crispr_muts$Mutation_Status <- ifelse(is.na(crispr_muts$Mutation_Status), 0, crispr_muts$Mutation_Status)
## Identify point mutations and wildtypes for testing
crispr_muts$PointMutation <- ifelse(is.na(crispr_muts$PointMutation), 1, crispr_muts$PointMutation)
# Summarize number of mutant and Wildtype cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Gene) %>%
  summarize(N_Wildtype = length(Mutation_Status) - sum(Mutation_Status == 1),
            N_Mutant = sum(Mutation_Status == 1))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Gene")
# Add Color columns
crispr_data$Color <- ifelse(crispr_data$Mutation_Status == 0, "thistle", "darkorchid")
crispr_data$Color <- factor(crispr_data$Color)

# Cell line lineages
# crispr_data <- merge(crispr_data, ccl_converter, by = c("CCLE_Name", "Broad_ID"), all.x = TRUE)
# levels(crispr_data$lineage_name) <- sort(levels(crispr_data$lineage_name), decreasing = TRUE)
# Copy number
crispr_data <- merge(crispr_data, cn_melt, by = c("Gene", "Broad_ID"), all.x = TRUE)
# Gene expression (RPKM)
ge_crispr <- filter(ge_melt, Gene %in% as.character(unique(crispr_melt$Gene)))
# ge_crispr_19Q1 <- filter(ge_melt_19Q1, Gene %in% as.character(unique(crispr_melt$Gene)))
crispr_data <- merge(crispr_data, ge_crispr, by = c("Gene", "Broad_ID"), all.x = TRUE)
# crispr_data <- merge(crispr_data, ge_crispr_19Q1, by = c("Gene", "Broad_ID"), all.x = TRUE)
saveRDS(crispr_data, "./../htscreens_giant_files/crispr_data_19Q1.rds", compress = "xz")
```

@@

# Filter G2P CRISPR for point mutations
crispr_data_ptmuts <- filter(crispr_data, PointMutation == 1)
crispr_data_ptmuts <- subset(crispr_data_ptmuts, select = -c(PointMutation))
# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Gene %in% g2p_genes)
crispr_data_g2p <- unique(crispr_data_g2p)
crispr_data_g2p$Gene <- factor(crispr_data_g2p$Gene, levels = g2p_genes)
crispr_data_g2p$Mutation_Status <- factor(crispr_data_g2p$Mutation_Status)
```

>>>>

MergeOmicsG2P <- function(database, outfile) {
  # Merge mutation status, CN, and GE omics data with drugscreen results and metadata
  filt <- auc %>% filter(Database == database) %>% select(Cell_line, Histology, TCGA_classification, CCLE_Name, Broad_ID, Drug, AUC_IC50)
  print(paste0("Filtered for ", database, "."))
  colnames(filt)[colnames(filt) == "AUC_IC50"] <- "Score"

@@

  # Merge mutation calls
  data <- merge(data, maf_df_filt, by = c("Gene", "Broad_ID"), all.x = TRUE)
  data$Mutation_Status <- ifelse(is.na(data$Mutation_Status), 0, data$Mutation_Status)
  ## Identify point mutations
  data$PointMutation <- ifelse(is.na(data$PointMutation), 1, data$PointMutation)
  ## Add color columns for plotting
  if(database == "CCLE") {
    data$Color <- ifelse(data$Mutation_Status == 0, "palegreen3", "springgreen4")
    data$Color <- factor(data$Color)
  }
  if(database == "CTRP") {
    data$Color <- ifelse(data$Mutation_Status == 0, "slategray3", "steelblue4")
    data$Color <- factor(data$Color)
  }
  if(database == "GDSC") {
    data$Color <- ifelse(data$Mutation_Status == 0, "paleturquoise3", "turquoise4")
    data$Color <- factor(data$Color)
  }
  print("Merged mutation calls.")

@@

  # Drop cell lines not in DepMap
  data <- filter(data, !is.na(Broad_ID))

@@

FilterForPointMuts <- function(dataframe, outfile) {
  # Filter for point mutations
  ptmuts <- filter(dataframe, PointMutation == 1)
  # Remove columns that lead to duplicate mutation entries
  ptmuts <- subset(ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, PointMutation))
  ptmuts$Gene <- ordered(ptmuts$Gene, levels = g2p_genes)
  ptmuts <- unique(ptmuts)
  saveRDS(ptmuts, outfile, compress = "xz")
  return(ptmuts)
}

adj_signif <- function(df, alpha) {
  # Takes df from compare_means and creates a signif code column for FDR-adjusted p-vals
  df$p.signif.adj <- ifelse(df$p.adj <= alpha, "*", NA)
  df$p.signif <- ifelse(df$p.signif == "ns", NA, df$p.signif)
  df$p.short <- formatC(df$p, format = "g", digits =  2)
  df$p.adj.short <- formatC(df$p.adj, format = "g", digits =  2)
  return(df)
}

WilcoxonTestDF <- function(dataframe, measure_var, grouping, outfile) {
  require(ggpubr)
  signif <- compare_means(Score ~ Mutation_Status, group.by = grouping, data = dataframe, method = "wilcox.test", p.adjust.method = "BH")
  signif <- adj_signif(signif, alpha = 0.05)
  signif <- signif[order(signif$p),]
  saveRDS(signif, outfile)
  return(signif)
}
```

@@

ccle <- MergeOmicsG2P(database = "CCLE", outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp <- MergeOmicsG2P(database = "CTRP", outfile = "./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- MergeOmicsG2P(database = "GDSC", outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")

ccle_ptmuts <- FilterForPointMuts(dataframe = ccle, outfile = "./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
# ctrp_ptmuts <- FilterForPointMuts(dataframe = ctrp, outfile = "./data_munging/rds/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- FilterForPointMuts(dataframe = gdsc, outfile = "./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")
```

@@

```{r, eval=FALSE}
ccle <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
# ctrp <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes.rds")
gdsc <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
# ctrp_ptmuts <- readRDS("./../htscreens_giant_files/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

# Testing
test_ccle <- ccle_ptmuts %>% count(Gene, Mutation_Status)
# test_ctrp <- ctrp_ptmuts %>% count(Gene, Mutation_Status)
test_gdsc <- gdsc_ptmuts %>% count(Gene, Mutation_Status)
```

@@

```{r, eval=FALSE}
crispr <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
colnames(crispr)[1] <- "Broad_ID"

# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "Broad_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Gene", value.name = "Score")

# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_meta, by = "Broad_ID", all.x = TRUE)

# Merge mutation annotations
maf_filt_crispr <- filter(maf_df, Gene %in% as.character(unique(crispr_melt$Gene)))
maf_filt_crispr <- subset(maf_filt_crispr, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1))
maf_filt_crispr <- unique(maf_filt_crispr)
crispr_muts <- merge(crispr_melt, maf_filt_crispr, by = c("Gene","Broad_ID"), all.x = TRUE)
crispr_muts$Gene <- factor(crispr_muts$Gene)
crispr_muts$Mutation_Status <- ifelse(is.na(crispr_muts$Mutation_Status), 0, crispr_muts$Mutation_Status)
## Identify point mutations and wildtypes for testing
crispr_muts$PointMutation <- ifelse(is.na(crispr_muts$PointMutation), 1, crispr_muts$PointMutation)

# Summarize number of mutant and Wildtype cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Gene) %>% summarize(N_Wildtype = length(Mutation_Status) - sum(Mutation_Status == 1), N_Mutant = sum(Mutation_Status == 1))
## Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Gene")
## Add Color column
crispr_data$Color <- ifelse(crispr_data$Mutation_Status == 0, "thistle", "darkorchid")
crispr_data$Color <- factor(crispr_data$Color)

# Cell line lineages
crispr_data <- merge(crispr_data, cn_melt, by = c("Gene", "Broad_ID"), all.x = TRUE)

# Gene expression (RPKM)
ge_crispr <- filter(ge_melt, Gene %in% as.character(unique(crispr_melt$Gene)))
crispr_data <- merge(crispr_data, ge_crispr, by = c("Gene", "Broad_ID"), all.x = TRUE)
saveRDS(crispr_data, "./../htscreens_giant_files/crispr_data_19Q1.rds", compress = "xz")
```

@@

# Filter G2P CRISPR for point mutations
crispr_data_ptmuts <- filter(crispr_data, PointMutation == 1)
crispr_data_ptmuts <- subset(crispr_data_ptmuts, select = -c(PointMutation))
# Filter CRISPR for G2P genes in G2P
crispr_data_g2p <- filter(crispr_data_ptmuts, Gene %in% g2p_genes)
crispr_data_g2p <- unique(crispr_data_g2p)
crispr_data_g2p$Gene <- factor(crispr_data_g2p$Gene)
crispr_data_g2p$Mutation_Status <- factor(crispr_data_g2p$Mutation_Status)
crispr_data_g2p$Primary.Disease <- factor(crispr_data_g2p$Primary.Disease)
```

# WILCOXON TESTS

```{r, eval=FALSE}
crispr_signif_gene <- WilcoxonTestDF(dataframe = crispr_data_g2p, grouping = "Gene", outfile = "./data_munging/rds/crispr_signif_g2p_gene.rds")
crispr_signif_lineage <- WilcoxonTestDF(dataframe = crispr_data_g2p, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/crispr_signif_g2p_lineage.rds")

ccle_signif_gene <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = "Gene", outfile = "./data_munging/rds/ccle_signif_g2p_gene.rds")
ccle_signif_lineage <- WilcoxonTestDF(dataframe = ccle_ptmuts, grouping = c("Gene", "Primary.Disease"), outfile = "./data_munging/rds/ccle_signif_g2p_lineage.rds")
```

# SESSION INFORMATION

```{r}
print(sessionInfo())
```
